Equivalence Testing and the Second Generation P-Value

### 2018-07-18 sgpv_vs_TOST.Rmd
new TOST_to_SGPV file that uses one TOSTone.raw function output
add plot

---
title: "The Second Generation P-Value is Old Wine in New Bottles"
author: "Daniël Lakens"
date: "1 juli 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

  # special case 2: point H0 & non-point CI
  # pdelta = 1/2 if H0 is inside the CI
  # pdelta = 0 if H0 is outside the CI
  if (delta_lb == delta_ub & lb != ub) {
    if (delta_lb <= ub & delta_lb >= lb) {
      return(1/2)
    } else {
      return(0)
    }
  }
}
```

@@

res <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)

p_tost <- res$TOST_p2
p_delta(mu + res$LL_CI_TTEST,
        mu + res$UL_CI_TTEST, 
        mu + low_eqbound, 
        mu + high_eqbound)

#SGPV
ll_ci <- (m - mu)-(1.96*(sd/sqrt(n)))
ul_ci <- (m - mu)+(1.96*(sd/sqrt(n)))

(min(ul_ci, high_eqbound) - max(ll_ci, low_eqbound)) / min(ul_ci - ll_ci, 2 * (high_eqbound - low_eqbound))

#The TOST p-value is calculated as:

t <- (m - mu - low_eqbound)/(sd/sqrt(n))
p <- pt(t, n-1, lower.tail = FALSE)
p

pTOST_converter(p_tost=res$TOST_p2,
                bound = "high",
                sd = sd,
                n = n,
                alpha = alpha,
                low_eqbound = low_eqbound,
                high_eqbound = high_eqbound)

@@

It should be possible to compute how these two are directly related, to calculate the one in the other! But how?

>>>>

---
title: "Equivalence Testing and the Second Generation P-Value"
author: "Daniël Lakens & Marie Delacre"
date: "1 juli 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

  # special case 2: point H0 & non-point CI
  # pdelta = 1/2 if H0 is inside the CI
  # pdelta = 0 if H0 is outside the CI
  if (delta_lb == delta_ub & lb != ub) {
    if (delta_lb <= ub & delta_lb >= lb) {
      return(1/2)
    } else {
      return(0)
    }
  }
}

#Code to calculate SGPV fron TOST results
# Input parameters
# res = the stored results from the TOST analysis
# n is the sample size
# sd is the standard deviation
#-----------------------------------------------------------------------
TOST_to_SGPV <- function(tost_res){

  # computing t, based on p_tost
  t <- min(tost_res$TOST_t1, tost_res$TOST_t2)

  n <- tost_res$TOST_df + 1
  sd <- (tost_res$diff-tost_res$low_eqbound)/tost_res$TOST_t1 * sqrt(n)

  # computing I
  dist <- t * sd/sqrt(n) # how far is I from the closest eqbound?

  #LAKENS: I I fixed this now - you can't know based on non-sig if bound is left or right from alpha. 
  #It should be based on whether p is larger or smaller than 0.5. I smaller than 0.5, bound - dist. 
  if (tost_res$TOST_p1 > tost_res$TOST_p2) {
    if (tost_res$TOST_p1 < 0.5){I = tost_res$low_eqbound + dist}
    if (tost_res$TOST_p1 > 0.5){I = tost_res$low_eqbound - dist}}

  if (tost_res$TOST_p2 > tost_res$TOST_p1) {
    if (tost_res$TOST_p2 < 0.5){I = tost_res$low_eqbound - dist}
    if (tost_res$TOST_p2 > 0.5){I = tost_res$low_eqbound + dist}}

  # Computing the CI around I
  lb <- I - qt(1 - tost_res$alpha/2, df = n - 1) * sd/sqrt(n) # lower bound
  ub <- I + qt(1 - tost_res$alpha/2, df = n - 1) * sd/sqrt(n) # upper bound

  if (lb > tost_res$high_eqbound | ub < tost_res$low_eqbound){
    SGPV = 0
    } 
  else if(lb > tost_res$low_eqbound & ub < tost_res$high_eqbound){
    SGPV = 1
    } 
  else {SGPV <- (min(ub, tost_res$high_eqbound) - max(lb, tost_res$low_eqbound))/min(ub - lb, 2 * (tost_res$high_eqbound - tost_res$low_eqbound))}

  return(SGPV)
}

@@

tost_res <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)

TOST_to_SGPV(tost_res = tost_res)

@@

The confidence interval width is a uniformly distributed across the mean differences, in the sense that as the true mean in a one-sample t-test gets closer to the test value (in the plot below, from situation A to D, the mean gets closer to the test value by 0.1) the difference in the overlap is stable.

tost_res1 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV1 <- TOST_to_SGPV(tost_res = tost_res1)

m <- 145.9
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

tost_res2 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV2 <- TOST_to_SGPV(tost_res = tost_res2)

m <- 145.8
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

tost_res3 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV3 <- TOST_to_SGPV(tost_res = tost_res3)

m <- 145.7
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

tost_res4 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV4 <- TOST_to_SGPV(tost_res = tost_res4)

0.7551064 - 0.7040851
0.6530639 - 0.6020426

plot(NA, 
     ylim = c(0, 1), 
     xlim = c(-3, 3),
     yaxt = "n",
     ylab = "",
     xlab = "Mean Difference")
axis(2, at = c(0.2,0.4,0.6,0.8), labels = c("D", "C", "B", "A"), las = 1)
abline(v = tost_res1$high_eqbound, 
       lty = 2)
abline(v = tost_res1$low_eqbound, 
       lty = 2)
abline(v = 0, 
       lty = 2, 
       col = "grey")

points(x = tost_res4$diff, 
       y = 0.2, 
       pch = 15, 
       cex = 2)
segments(tost_res4$LL_CI_TOST, 
         0.2, 
         tost_res4$UL_CI_TOST, 
         0.2, 
         lwd = 3)

points(x = tost_res3$diff, 
       y = 0.4, 
       pch = 15, 
       cex = 2)
segments(tost_res3$LL_CI_TOST, 
         0.4, 
         tost_res3$UL_CI_TOST, 
         0.4, 
         lwd = 3)

points(x = tost_res2$diff, 
       y = 0.6, 
       pch = 15, 
       cex = 2)
segments(tost_res2$LL_CI_TOST, 
         0.6, 
         tost_res2$UL_CI_TOST, 
         0.6, 
         lwd = 3)

points(x = tost_res1$diff, 
       y = 0.8, 
       pch = 15, 
       cex = 2)
segments(tost_res1$LL_CI_TOST, 
         0.8, 
         tost_res1$UL_CI_TOST, 
         0.8, 
         lwd = 3)

For example, the SGPV from A to D is `r SGPV1`, `r SGPV2`, `r SGPV3`, and `r SGPV4`. The difference in the percentage of overlap between A and B (`r SGPV1-SGPV2`) is identical to the difference in the percentage of overlap between C and D as the mean gets 0.1 closer to the test value (`r SGPV3-SGPV4`).
